### 2018-02-13 response_letter.txt
Add 'response letter'

----------------------- REVIEW 1 ---------------------
PAPER: 139
TITLE: Fixed-Parameter Tractable Sampling for RNA Design with Multiple
Target Structures
AUTHORS: Stefan Hammer, Yann Ponty, Wei Wang and Sebastian Will

----------- Overall evaluation -----------
The paper presents a framework based on a basic method in
Parameterized Complexity to handle some aspects, which are argued to
be computational bottlenecks, of the design of multi-stable RNA
molecules. The central problem formulated in this paper concerns the
computation of a partition function over sequences. The precise
definition of this problem was not at all clear to me when I read the
paper for the first time (see page 3), given that there seems to be
overload of notations (in particular, S is given as argument but also
appears in the sum). Generally, the definitions given in this paper
are mostly intuitive/vague rather than mathematically
precise. However, I did understand the problems that are actually
solved subsequently from their specific explanations and
solutions. Thus, the intent of the authors can be deciphered if
desired. Given that the paper is not submitted to a TCS conference, it
might be that this writing style is in fact more appropriate, and
hence I do not know whether it should be criticized.

Firstly, the authors consider a problem where we are given a graph, and we
need to assign a letter (A,G,C or U) to each vertex so that every two
adjacent vertices u,v are assigned one of the following configurations:
{A,C}, {G,C} or {G,U}. The authors show that the problem of counting the
number of solutions to this problem is #P-hard by a simple reduction from
Independent Set on bipartite graphs.

Secondly, the authors present an algorithm for sampling
Boltzmann-weighted designs. Here, they solve a more general problem
than only the counting problem above, but can compute a partition
function depending on the assignments. The result is a parameterized
algorithm with respect to the treewidth of the input graph (Section
4.1). The design of this algorithm is completely straightforward. The
authors further suggest an optimization of their algorithm that is
based on a simple yet nice observation (from a previous paper) that
speeds-up the computation. The authors use their own notation rather
than standard notation in Parameterized Complexity. In general, the
presentation and proofs of the algorithms are quite cumbersome, making
them appear more difficult than they truly are. There also seems to be
several typos in the presentation (e.g., how can we have a|a' in the
entry m_{w\rightarrow u}?). Nevertheless, the ideas of the algorithms
themselves are correct. Next, the computation is further extended by
enriching the functions computed at individual nodes to suit
multidimensional sampling.

The algorithms above were implemented, and the authors conducted several
experiments to evaluate their approach. I have carefully read this
section, but given that this subject is completely outside my area of
expertise, I cannot comment on this. It seems to me that this part has
taken most of the efforts invested in this project. The authors also argue
here that in practice, the treewidth of the graphs relevant to their
applications will be small.

Overall, the acceptance of this paper depends on whether the biological
motivation behind it is reasonable and of interest. The technique used is
a straightforward application of a known method in Parameterized
Complexity (DP over tree decomposition for counting), and the hardness
proof is also quite simple. However, I greatly appreciate seeing this
technique applied in a new context, as well as seeing such algorithms
implemented and used in practice. In this sense, the paper is indeed quite
novel. I therefore believe that, if the biological motivation is good
enough, this paper can greatly advance relations between bioinformatics
and parameterized analysis. Although I disliked the presentation (I would
have preferred much more formal, precise definitions and arguments,
preferably also using standard notations), I have carefully read the
proofs themselves and the results in the paper are correct. I therefore
recommend acceptance, but with very low confidence.

-------------------------------------------------------------------------------------------
Main Reviewer's comment: My sub-reviewer is an expert in parameterized
complexity rather than bioinformatics. She originally set the confidence
to 2, stating that it is up to the other reviewers to decide if the
application is important in bioinformatics. My opinion is that the
application (sampling for multi-target RNA design) is an important, highly
motivated problem in structural RNAomics. Furthermore, the fact that the
authors implemented their algorithm as a tool and made their code
available via GitHub is appreciated.  Therefore, I raised the confidence
score.

----------------------- REVIEW 2 ---------------------
PAPER: 139
TITLE: Fixed-Parameter Tractable Sampling for RNA Design with Multiple
Target Structures
AUTHORS: Stefan Hammer, Yann Ponty, Wei Wang and Sebastian Will

----------- Overall evaluation -----------
The paper presents an algorithm for sampling RNA sequences with the
aim of obtaining a sequence whose energy in a given set of secondary
structures is close to a desired vector of target energies.

The algorithm is based on dynamic programming on a tree decomposition
of a graph that is built from the constraints by the given secondary
structures. The main and most difficult part of the algorithm is to
fill a dynamic programming table for computing the partition functions
of the given energies. Afterwards, the sampling is performed using the
entries of the dynamic programming table.

In addition to this algorithm a negative result about the computation
of the number of valid RNA sequences for a given set of secondary
structures is given, stating that this number cannot be computed in
polynomial time unless #P=P.

The approach to use dynamic programming on a dependency graph in the
context of sampling seems novel. The algorithm itself works more or
less as expected. The authors describe one improvement over brute-force DP
that is based on the fact that the constraint graph obtained from the
structures is bipartite.

The experimental part of the paper demonstrates that the sampling
based on the new approach leads to better results than uniform
sampling both in terms of the energy distribution of the sample
population itself but also if the samples are used as starting
solutions in a local search algorithm. A small caveat is that the
paper does not compare with any of the previous approaches for RNA
Design but in my opinion, the sampling procedure itself is
sufficiently important to make the experiments meaningful.

Overall, I recommend to accept the paper. The topic and the treatment
with dependency networks make it a timely contribution. The paper is
reasonably well-written.

Specific suggestions for improvements:

-Page 1: "The overall complexity ... grows like Theta(4^gamma) where gamma
is a parameter whose value is bounded only by the length" This seems
rather strange because if gamma is the overall length, then 4^gamma can be
trivially obtained by enumerating all sequences and then computing the
energies for each sequence. Maybe it is supposed to mean that in the
worst-case gamma can be the length of the sequence? But this is also true
for the treewidth making this remark maybe a bit unfair. I suggest to be
more precise and explain roughly what gamma is.

>>>>> Elaborate that gamma=n is worst case and in practice much smaller due to ear decomp (similar to tree width)

-Page 2 "in a principled way" Could you provide some further explanation
what "principled" means here?

>>>>> --> YANN; something like 'generic'?

-Page 3 The definition of RNA structure allows that one position i takes
part in more than one base pair (e.g. {(1,2),(1,3)} is a set of base pairs
and thus a secondary structure according to the definition).

>>>>> add this restriction (we do not need degree>1, even if it does not harm)

-Page 3 "{A,C}" -> "{A,U}"

-Page 3 "(Si,Sj)" -> "{Si,Sj}"

-Page 3 Please highlight a bit better when the compatability graph is of
interest and when the dependency graph is of interest. It seems that the
compatability graph is only interesting in Section 3?

-Page 3 It seems strange that the partition function depends on S while S
is used in the sum. Should it be "Z_E_R(n)" or "Z_E_R(S_n)"?

>>>> Embarassing typo: this should be Z_E_R(\R)

-Page 3 "as the minimal set" Is it necessarily unique? If not, "as a
minimal set".

-Page 3 "an hypergraph" -> "a hypergraph"

-Page 4 "base pair models" -> "base pair model"

-Page 4 "ins polynomial time" -> "in polynomial time"

-Page 4 "strongly connected" Why "strongly" is the graph directed? There
is no discussion of weak vs. strong connected components.

-Page 4,5 two missing "\qed"s

-Page 5 In the definition of "f[a]" one needs that "dep(f)" contains "dom(A)"

>>>> CHECK

-Page 5 The difference variables should probably be defined only for
adjacent bags of the tree decomposition? Otherwise, I don't see how the
claim about the one-variable difference is true.

-Page 5 "Denote by A(Y)" What is Y?

-Page 6 In Line 6 of the algorithm, I don't see how m_w->u has an entry
for "[a|a']" as "a" may contain some new variables.

-Page 6 "Its correctness" -> "The correctness of the algorithm"

-Page 6 Running times are not reported here but it seems crucial to obtain
good tree decompositions, I am not sure whether greedy fill-in is the best
strategy. A more recent approach is for example Flow-Cutter by Ben
Strasser (github.com/kit-algo/flow-cutter-pace17).

-Page 6 "requires O(..." + "time"

-Page 7 For Theorem 1, I was wondering whether it is really necessary to
consider the total number of connected components. It seems that only the
maximum number of components per bag is crucial?

-Page 7 "we encode the network energy ..." I was a bit confused here. What
is w_l? Is it possible to give the formula?

-Page 8 Should it "ln" instead of "log"?

>>>>> YES

-Page 8 What is "gamma" in "gamma^{mu_l}"?

-Page 8 "expected number grows like Theta(1)" This sounds a bit awkward as
Theta(1) doesn't really grow. How about "expected number is Theta(1)"?

>>>> REFORMULATE

-Page 9 In the definition of f(S), the first sum should be from l=1?

-Page 14 Lemma 1 is a well-known fact about tree decompositions. It does
not need a proof and should be found in some surveys about treewidth.

>>>>> AGREED (Discuss with Yann: drop proof and cite some text book)

----------------------- REVIEW 3 ---------------------
PAPER: 139
TITLE: Fixed-Parameter Tractable Sampling for RNA Design with Multiple
Target Structures
AUTHORS: Stefan Hammer, Yann Ponty, Wei Wang and Sebastian Will

----------- Overall evaluation -----------
The manuscript by Hammer et al. describes an general theoretical framework
for designing RNA sequences with multiple target secondary structures.
First, the authors show that the problem of counting the number of
possible designs is #P-hard. Then, they develop an algorithm for
calculating the partition function, and sampling design instances from
this distribution. The complexity of these algorithms is determined by the
tree-width of the constraint graph.

The method is timely and nicely expands recent contributions in the same
area. Moreover, I believe that the nature of the work and theoretical
results presented in this paper should be of interest for the broad RECOMB
community.

Comments:

* There are a couple of shortcuts in the notations that occasionally
distracted me (although it was correct). For instance, in definition 1,
point 3, e is an edge and X(v) is a set of vertice. Or in the definition
of E_R(S) the dependency between F and R is not explicited. Similarly,
definition 2 introduces the RNA design networks but the definition stays
at a relatively abstract level. If the object is associated with RNA, why
not using a specialized terminology (e.g. nucleotide instead of
variables)?

>>>>> Notation could be adapted; e.g. "nucleotide instead of
variables" would work for this paper (but not for all possible extensions)

* The experimental validation is limited, but it is (to my point-of-view)
not an issue as the core of the paper are the theoretical results. If
possible, it would be useful to expand this benchmark. For instance, by
including a discussion using known multi-stable RNAs and comparing real
and artificial sequences. It would be interesting to characterize which
parameters increase the difficulty of the design, or determine a priori
the range of possible target energies.

>>>>> Don't we already have artificial vs. real; can we write some discussion?

* This is an open question, but I was wondering how the authors could use
their algorithm to control the energy barrier between two target
structures. For instance, one could identify a structure near the saddle
point and force the design to assign a high energy to the latter (and by
constrast low energies to the targets).

Controling the energy barrier between two stable structure An important
feature when designing multi-stable RNA

>>>>> INTERESTING IDEA; rather for design than sampling. We could try to optimize 
the direct path barrier in this way (for producing decoys...)

* It was not clear to me if the authors implemented dangling ends?

>>>>> NO; if possible, reformulate the resepctive sentence

Typos:
* Eq 1: Z is a function of S but S is in the sum as well...
* Sec. 2, start of paragraph 2: "Given is"?