### 2018-09-19 RNARedPrint.tex
* Added 'contribution' section to intro * Reshaped complexity paragraph

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
%\documentclass{bmcart}

%%% PACKAGES

\usepackage{xr}
\externaldocument{supplement}

\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
% \usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
%   \let\itemize\compactitem
%   \let\enditemize\endcompactitem
%   \let\enumerate\compactenum
%   \let\endenumerate\endcompactenum
%   \let\description\compactdesc
%   \let\enddescription\endcompactdesc
%   \pltopsep=1pt
%   \plitemsep=1pt
%   \plparsep=1pt
\usepackage{hyperref}
\usepackage{xspace}
%\usepackage{geometry}
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{longtable}
\usepackage[vlined]{algorithm2e}

\usepackage{inconsolata}

\usepackage{xifthen}
\usepackage{stmaryrd}

\usepackage{xcolor}

\usepackage{todonotes}

%\hypersetup{
%    bookmarks=false,         % show bookmarks bar?
%    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
%    pdftoolbar=true,        % show Acrobat’s toolbar?
%    pdfmenubar=true,        % show Acrobat’s menu?
%    pdffitwindow=true,     % window fit to page when opened
%    pdfstartview={FitH},    % fits the width of the page to the window
%    colorlinks=true,       % false: boxed links; true: colored links
%    linkcolor=red!70!black,          % color of internal links (change box color with linkbordercolor)
%    citecolor=red!70!black,        % color of links to bibliography
%    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
%    citecolor=black,        % color of links to bibliography
%    filecolor=magenta,      % color of file links
%    urlcolor=red!70!black           % color of external links
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%TODO uncomment
%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Rolf's includegraphicstop
\makeatletter
\newsavebox{\@alignepsbox}
\newlength{\@aligneps}
\newcommand{\includegraphicstop}[2][]{%
\sbox{\@alignepsbox}{\includegraphics[#1]{#2}}%
\setlength{\@aligneps}{-\ht\@alignepsbox}%
\addtolength{\@aligneps}{2ex}%
\raisebox{\@aligneps}{\usebox{\@alignepsbox}}}
\makeatother

%\makeatletter
%\let\oldlt\longtable
%\let\endoldlt\endlongtable
%\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
%\def\longtable@i[#1]{\begin{figure}[t]
%\onecolumn
%\begin{minipage}{0.5\textwidth}
%\oldlt[#1]
%}
%\def\longtable@ii{\begin{figure}[t]
%\onecolumn
%\begin{minipage}{0.5\textwidth}
%\oldlt
%}
%\def\endlongtable{\endoldlt
%\end{minipage}
%\twocolumn
%\end{figure}}
%\makeatother

%%%%%%%%%%%%%%%%% Theorems %%%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{observation}[theorem]{Observation}

%\newtheorem{algorithm}{Algorithm}
\newtheorem{axiom}{Axiom}
\newtheorem{hypothesis}{Working Hypothesis}
\newenvironment{proof}[1][]{\noindent Proof\ifthenelse{\equal{#1}{}}{}{ (#1)}:~}{}

%%% macros for notation in DP framework
\newcommand{\network}{\mathcal{N}}
\newcommand{\val}{\bar S} % valuation aka assignment
\newcommand{\dep}{\operatorname{dep}}
\newcommand{\energy}[1]{\operatorname{e}_{#1}}
\newcommand{\numberof}{\operatorname{\#}}
\newcommand{\partfun}[1]{Z_{#1}}
\newcommand{\separator}[2]{\operatorname{sep}(#1,#2)}
\newcommand{\difference}[2]{\operatorname{diff}(#1 \rightarrow #2)}
\newcommand{\real}{\mathbb{R}}
\newcommand{\genmarg}[1]{(\!|\!#1\!|\!)}
\newcommand{\gencomb}[1]{\langle\!|#1|\!\rangle}
\newcommand{\Message}[2]{m_{#1\rightarrow #2}}

\newcommand{\energyModel}{{\cal M}}
\newcommand{\structureElements}{{\cal SE}}
\newcommand{\powerSet}[1]{2^{#1}}
\newcommand{\underConstruction}[1]{{\LARGE$\triangle$\Large\!\!\!\!!}$\quad$\textcolor{red}{#1}}
\newcommand{\argmin}{\operatorname*{arg\,min}}
\newcommand{\objective}{{\mathbb{F}}}

\newcommand{\partseqs}{\mathcal{P\!S}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\R}{\mathcal{R}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}

\newcommand{\width}{w}

\newcommand{\sample}{\texttt{Sample}}
\newcommand{\elim}[2]{\operatorname{elim}(#1,#2)}

%\newcommand{\Ehp}[1]{E^{\textrm{hp}}(#1)}
%\newcommand{\Eint}[1]{E^{\textrm{int}}(#1)}

\newcommand{\EbpSym}{E^{\textrm{bp}}}
\newcommand{\Ebp}[2]{\EbpSym_{#1}(#2)}

\newcommand{\Def}[1]{\emph{#1}}

\newcommand{\TargetE}{E^{\star}}
%\newcommand{\MFE}{\text{\rm MFE}}
\newcommand{\EnsE}{\text{\rm G}} %ensemble energy

\newcommand{\Obj}{\text{\rm MultiDefect}}

\newcommand{\TODO}[1]{\textcolor{red!70!black}{\textbf{TODO: #1}}}

\newcommand{\parHead}[1]{\Final{\paragraph{#1}}}

\newcommand{\Final}[1]{\begingroup\color{red!70!black}#1\endgroup}
%% Uncomment the line below for ``Final'' version
\renewcommand{\Final}[1]{}

\newcommand{\Design}[1]{{\sf Designs}^{\star}(#1)}
\newcommand{\NumDesign}{\ensuremath{\#}{\sf Designs}\xspace}
\newcommand{\IS}[1]{{\sf IndSets}(#1)}
\newcommand{\Nuc}[1]{{\sf #1}}
\newcommand{\Ab}{\Nuc{A}}
\newcommand{\Cb}{\Nuc{C}}
\newcommand{\Gb}{\Nuc{G}}
\newcommand{\Ub}{\Nuc{U}}

\newcommand{\Objectiv}{\Nuc{U}}

\newcommand{\GCb}{\Gb\Cb}

\newcommand{\Software}[1]{{\ttfamily #1}}

\newcommand{\substitute}[2]{#1\cup#2}
%\newcommand{\evalfor}[2]{#1\llbracket{}#2\rrbracket{}}
\newcommand{\evalfor}[2]{#1(#2)}

\renewcommand{\gets}{:=}

\setlength{\parskip}{.2em}

\newcommand{\RNAblueprint}{{\tt \bfseries{}\color{black!85} RNA\textcolor{blue!70!black}{Blue}Print}}
\newcommand{\ourprog}{{\tt \bfseries{}\color{black!85}RNA\textcolor{red!70!black}{Red}Print}}

\newcommand{\Energy}[2]{E(#2;#1)}

\newcommand{\EnergyTurner}{E_{\text{T}}}
\newcommand{\EnergyStacking}{E_{\text{st}}}

\newcommand{\citep}[1]{\cite{#1}}
\newcommand{\citet}[1]{\cite{#1}}

%%% end macro defs

%% fix some weirdness of the bioinfo class (footer overlapping text on page 1)
\makeatletter
\setlength\footskip{30\p@}          % space above footer line
\makeatother

\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Methodology}
%\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%[Efficient Sampling for Multi-Target RNA Design]
\title{Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
    addressref={aff1,aff2,aff3},
    email={stefan.hammer@uni-leipzig.de}
]{\inits{SH} \fnm{Stefan} \snm{Hammer}}
%
\author[
    addressref={aff4},
    email={wei.wang@lix.polytechnique.fr}
]{\inits{WW} \fnm{Wei} \snm{Wang}}
%
\author[
    addressref={aff2,aff3},
    email={will@tbi.univie.ac.at}
]{\inits{SW} \fnm{Sebastian} \snm{Will$^*$}}
%
\author[
    addressref={aff4,aff5},
    corref={aff4},                       % id of corresponding address, if any
    email={yann.ponty@lix.polytechnique.fr}
]{\inits{YP} \fnm{Yann} \snm{Ponty}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \address{$^{\text{\sf 1}}$University Leipzig, Department of Computer Science and Interdisciplinary Center for Bioinformatics, 04107 Leipzig, Germany;
\address[id=aff1]{%                           % unique id
  \orgname{Dept.\ Computer Science, and Interdisciplinary Center
      for Bioinformatics, Univ.\ Leipzig},
  \street{H{\"a}rtelstr.\ 16-18},
  \postcode{D-04107},
  \city{Leipzig}, 
  \cny{Germany}
}

% $^{\text{\sf 2}}$University of Vienna, Faculty of Chemistry, Department of Theoretical Chemistry, 1090 Vienna, Austria;
\address[id=aff2]{%
    \orgname{Dept.\ Theoretical Chemistry, Univ.\ Vienna},
    \street{W{\"a}hringerstr.\ 17},
    \postcode{A-1090}
    \city{Wien},
    \cny{Austria}
}

% $^{\text{\sf 3}}$University of Vienna, Faculty of Computer Science, Research Group Bioinformatics and
% Computational Biology, 1090 Vienna, Austria;
\address[id=aff3]{%
    \orgname{Bioinformatics and Computational Biology Research Group, Univ.\ Vienna},
    \street{W{\"a}hringerstr.\ 17},
    \postcode{A-1090}
    \city{Wien},
    \cny{Austria}
}

% $^{\text{\sf 4}}$CNRS UMR 7161 LIX, Ecole Polytechnique, Bat. Turing, 91120 Palaiseau, France;
\address[id=aff4]{%
    \orgname{CNRS UMR 7161 LIX, Ecole Polytechnique},
    \street{Bat. Alan Turing},
    \postcode{91120}
    \city{Palaiseau},
    \cny{France}
}

% $^{\text{\sf 5}}$ AMIBio team, Inria Saclay, Bat Alan Turing, 91120 Palaiseau, France}
\address[id=aff5]{%
    \orgname{AMIBio team, Inria Saclay},
    \street{Bat. Alan Turing},
    \postcode{91120}
    \city{Palaiseau},
    \cny{France}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}
\begin{abstract}
  \parttitle{Background}
  The design of multi-stable RNA molecules has important applications in biology, medicine, and biotechnology. Synthetic design approaches profit strongly from effective in-silico methods, which impact their cost and feasibility. \\
%
  \parttitle{Results} We devise a novel approach to a central ingredient of most in-silico
  design methods: the generation of sequences with multiple structural targets. For this task, we present the efficient (fixed-parameter tractable) approach \ourprog{}.
  % 
  Modeling
  the problem by specializing constraint networks, \ourprog{} supports generic
  Boltzmann-weighted sampling, which enables the positive design of RNA sequences
  with specific free energies (for each of multiple, possibly pseudoknotted, target structures) and \GCb-content.
%
  Securing the theoretical foundations of our method, we establish the $\#${\sf P}-hardness of uniform sampling. 
%
  Finally, we empirically study general properties of the approach and generate
  biologically relevant multi-target Boltzmann-weighted designs for a
  common design benchmark, including pseudoknots.
  Generating seed sequences with \ourprog{}, we demonstrate significant improvements over the previously best multi-target sampling strategy.\\
%
  \parttitle{Conclusion} \ourprog{} introduces a novel, flexible and---as we demonstrate--effective approach to multi-target RNA design, which promises broad applicability and extensibility.

  \smallskip
  \noindent
  Our free software is available at: \url{https://github.com/yannponty/RNARedPrint}\\

  \noindent
  Supplementary data are available online.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
  \kwd{RNA multi-target design}
  \kwd{RNA secondary structure}
  \kwd{Multi-dimensional Boltzmann sampling}
  \kwd{{$\#$}{\sf P}-hardness of RNA design}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
\parHead{Design, applications and motivation for multiple design.}Synthetic biology strives for the engineering of artificial biological
systems, promising broad applications in biology, biotechnology and
medicine. Centrally, this requires the design of biological
macromolecules with highly specific properties and programmable functions.
RNAs are particularly well-suited tools for
rational design targeting specific functions~\citep{Kushwaha2016}: on the one hand, RNA function is tightly
coupled to the formation of secondary structure, as well as changes in
base pairing propensities and the accessibility of regions, e.g.\ by
burying or exposing interaction sites~\citep{Rodrigo2014}; on the other hand, the
thermodynamics of RNA secondary structure is well understood and its prediction is
computationally tractable~\citep{McCaskill1990}. Thus, in rational design approaches, structure can serve as effective
proxy for, the ultimately targeted, catalytic or regulatory functions~\citep{Zhang2013}.

\parHead{Motivating multiple RNA design.} The function of many RNAs
depends on their selective folding into one or several alternative
conformations. Classic examples include riboswitches, which
adopt different stable structures upon binding a specific
ligand. Riboswitches have been a popular application of rational
design~\citep{Wachsmuth2013,Domin2017}, partly motivated by their
capacity to act as biosensors~\citep{Findeiss2017}.
In particular due to the kinetic coupling of RNA folding with RNA transcription, RNA families can feature alternative,
evolutionarily conserved, transient structures~\citep{Zhu2013}, which
are essential for the formation of their functional structures.  More generally, simultaneous compatibility to multiple
structures is a relevant design objective for engineering kinetically
controlled RNAs, finally targeting prescribed folding pathways. Thus,
advanced applications of RNA design often target multiple structures,
additionally aiming at other features, such as specific
\GCb-content (\GCb\%) as previously done by~\citet{Reinharz2013} or the presence/absence of
functionally relevant motifs (either anywhere or at specific
positions)~\citep{Zhou2013}; these objectives motivate flexible
computational design methods.

\parHead{On the importance of sampling for design.}
Many computational methods for RNA design follow the ``generate-and-optimize'' strategy: they randomly generate one or several \Def{seed} sequences, which are then optimized.
In many cases, the seed quality was
found to be critical for the empirical performance of RNA design
methods~\citep{Levin2012}. Generally, random seed generation
improves the prospect of subsequent optimizations, helping to overcome
local optima of the objective function, and increases the diversity
across designs~\citep{Reinharz2013}.  For single-target approaches,
\Software{INFO-RNA}~\citep{Busch2006} could significantly improve the success rate over \Software{RNAinverse}~\citep{Hofacker1994}, by starting its local search from the minimum energy sequence
for the target structure. Since this
strategy typically designs sequences with unrealistically high
\GCb\%,
more recent approaches like \Software{antaRNA}~\citep{Kleinkauf2015} and \Software{IncaRNAtion}~\citep{Reinharz2013} explicitly control \GCb\%; the latter applying adaptive sampling.

\parHead{Specificities and similarities of multi-target design.}
In the case of multi-target design, virtually all available methods~\citep{Lyngsoe2012,HoenerzuSiederdissen2013,Taneda2015,Hammer2017} follow the same overall generate-and-optimize strategy.
%
Facing the complex sequence constraints induced by multiple targets, early methods such as \Software{Frnakenstein}~\citep{Lyngsoe2012} and \Software{Modena}~\citep{Taneda2015} did not attempt to solve sequence generation systematically, but rely on ad-hoc sampling strategies.
%
Recently, the approach \Software{RNAdesign}~\citep{HoenerzuSiederdissen2013}, coupled with local
search within \RNAblueprint{}~\citep{Hammer2017}, solved the
problem of sampling seeds from the \emph{uniform} distribution for multiple
target structures. \Software{RNAdesign} adopts a graph coloring perspective,
coloring the sequence positions with nucleotides, such that the ends of each base pair receive compatible colors. Initially the method decomposes the graph hierarchically, using ear decomposition, and then \Def{precomputes} the number of valid sequences
within each subgraph. The decomposition is then reinterpreted as a
decision tree to perform \Def{stochastic backtracking}, inspired by
Ding and Lawrence~\citep{Ding2003}. Uniform sampling is achieved by
choosing individual nucleotide assignments with probabilities derived
from the subsolution counts. The overall complexity of
\Software{RNAdesign} grows like $\Theta(4^{\gamma})$, where the
parameter $\gamma$, while bounded only by the length of the designed RNA, is
typically much lower, due to the decomposition strategy.

\parHead{Motivation.}
While uniform sampling in \Software{RNAdesign}/\RNAblueprint{} turned out to be highly beneficial for general RNA design,
the work on these tools gave no definitive answer on the \emph{complexity of uniform sampling for multi-target design}. While time and space requirements grow exponential in the general case,
the decomposition strategy seems to allow practical applications.
%
Better understanding of the true complexity bounds is highly relevant for the application of such methods, in particular for applications in more expressive settings (e.g. Boltzmann sampling as discussed in this work or sampling satisfying multiple additional constraints and properties). All such extensions are still limited by the complexity of the multi-target uniform sampling problem, where counting of valid designs is the dominating step. In other words key questions include: \emph{How to sample, in a general way, from a Boltzmann distribution based on expressive energy models? How to enforce additional constraints, such as \GCb\%, complex sequence constraints, or the energy of individual structures?}

\subsection*{Contributions}

\begin{figure*}[t]
\begin{center}
    \includegraphics[width=.8\textwidth]{Figs/Workflow}
\end{center}
\caption{General outline of \ourprog{} for base pair-based energy models. From a set of target secondary structures (i), base-pairs are merged (ii) into a (base pair) dependency graph (iii) and transformed into a tree decomposition (iv). The tree is then used to compute the partition function, followed by a Boltzmann sampling of valid sequences (v). An adaptive scheme learns weights to achieve targeted energies and \GCb\%, leading to the production of suitable designs (vi).}
\label{fig:workflow}
\end{figure*}

Addressing these questions in more detail and generality leads to the practical advances in this work, which extend the expressivity, flexibility, and performance of multi-structure design. Namely, we introduce a generic framework (illustrated in~Fig.~\ref{fig:workflow}) enabling efficient Boltzmann-weighted sampling over RNA sequences with multiple target structures. As main algorithmic contribution of this work, we devise a dynamic programming algorithm, based on a tree decomposition, to compute partition functions and sample sequences from the Boltzmann distribution%
% (Subsection~\ref{sec:PF})
. We show that these algorithms are time and space-efficient, while requiring exponential time based on a specific property of the tree decomposition, called treewidth, whose value is typically low for typical instances; in short, our algorithms are fixed-parameter tractable (FPT).

This exponential dependency is arguably necessary, as the underlying counting problem is \#{\sf P}-hard, \emph{i.e.}  we show that there can not exists any efficient, polynomial time algorithm unless ${\sf P}={\sf NP}$. Our result relies on a surprising bijection between valid sequences and independent sets of a bipartite graph; the latter being the object of recent breakthroughs in approximate counting complexity~\citep{Bulatov2013,Cai2016}.

% Subsection~\ref{sec:complexity}
%
Due to the generality of our method, we can moreover strongly limit this parameter in practice by using state-of-the-art tree decomposition algorithms.
%Uniform sampling is handled as a special case of Boltzmann sampling, where each valid sequence receives energy zero and---consequently---computing partition functions specializes to counting.
By evaluating (partial) sequences in a specialized weighted constraint
network, we support arbitrary multi-ary constraints and thus
arbitrarily complex energy models,
notably subsuming all commonly
used RNA energy models%
%(Subsection~\ref{sec:energy_models})
.  Moreover, we describe an \Def{adaptive
  sampling} strategy to control the free energies of the individual
target structures and \GCb\%%
% (SubSection~\ref{sec:multiBoltzmann})
. %

We observe that
targeting realistic RNA energies in the
Turner RNA energy model
works well by performing sampling based on less complex RNA energy models, which keeps the treewidth low. This result is highly relevant in practice, since it allows to combine high efficiency with high accuracy and moreover leaves room for extensions of the method.

Eventually this strategy allows generating biologically relevant
multi-target designs, as we demonstrate by performing negative multi-target RNA design for a comprehensive, 
meaningful benchmark set~\citep{Taneda2015}.

\section*{Methods}
\subsection*{Problem statement}
\label{sec:problem-statement}
The main computational problem addressed in this work is to generate sequences over the alphabet $\Sigma=\{\Ab,\Cb,\Gb,\Ub\}$ with prescribed energies, for each single target structure, and $\GCb\%$.
Remarkably, such criteria can be homogeneously abstracted as constraints on the value of \Def{features}, each defined as sums over atomic feature contributions depending on limited subsets of sequence positions.

\paragraph{Central computational problem.}
Let us consider $k$ features $F_1,\dots F_k$ to target all of them simultaneously. We base our strategy for targeting the $k$ features $F_1,\dots,F_k$ on generating samples from a Boltzmann distribution parametrized by weights $\pi_1,\dots,\pi_k$, one for each feature. These weights essentially allow us to shift the sample distributions towards specific targets.
Sampling from this distribution requires to compute corresponding partition functions, such that we can draw sequences with probabilities proportional to their Boltzmann weight:
\begin{equation}
\label{eq:sample-distribution}
\Pr(S|\pi_1,\dots,\pi_k) \propto \prod_{1\leq \ell\leq k} \pi_\ell^{-F_\ell(S)}.
\end{equation}

The workhorse of our approach is the efficient (more precisely, fixed-parameter tractable) computation of feature-dependent partition functions over sequences.
Given are the features $F_1,\dots,F_k$, which encode the energies functions for the fixed structures and the function for the \GCb\%, all in terms of simple atomic energy contributions.
For weights $\pi_1,\dots,\pi_k$, one wants to compute the partition function
  \begin{equation}
    \label{eq:mainproblem}
    \partfun{\pi_1,\dots,\pi_k} = \sum_{S\in\Sigma^n} \prod_{1\leq \ell\leq k} \pi_\ell^{-F_\ell(S)}.
  \end{equation}

\paragraph{Representing \GCb\%-content and energies as features.}
Formally, we define a \Def{feature $F$} as function on sequences, whose value is obtained by summing over an associated set of \Def{atomic feature contributions}, briefly called \emph{contributions}. Each feature $f$ is associated with a restricted set of positions, its \emph{dependencies $\dep(f)$}, such that
$$
F(S) =\!\!\!\!\!\!\!\sum_{\substack{\text{$f$: $f$ is contribution of $F$,}\\\text{ $\dep(f)=\{x_1,\ldots,x_p\}$}}}\!\!\!\!\!\!\! f(\{x_1\mapsto S_{x_1},\ldots,x_p\mapsto S_{x_p}\}).
$$

Thus, the $\GCb\%$ feature is expressed by $n$ contributions $f^{\GCb}_i$ where $f^{\GCb}_i(\{i\mapsto S_i\})$ equals $-1$ if $S_i=\Gb$ or $\Cb$, and $0$ otherwise $(i=1,\dots,n)$.

For the definition of energy features, we abstract an \Def{RNA (secondary) structure $R$ of length $n$} as a set of \Def{base pairs} $(i,j)$, where
$1\leq i<j\leq n$ and for all different $(i,j), (i',j')\in R$: $\{i,j\}\cap\{i',j'\}=\emptyset$ (``degree $\leq$ 1'')---without precluding pseudoknots.
% We call an RNA structure $R$ \Def{non-crossing}, iff it does not
% contain any two different base pairs $(i,j)$ and $(i',j')$ such that
% $i\leq i'\leq j \leq j'.$

In a simple \Def{base pair energy model}, the energy of a target structure $R$ is expressed as feature with contributions $f^R_{ij}(\{i\mapsto x, j\mapsto y\})=\EbpSym(x,y)$ for each $(i,j)\in R,$ such that $F^R(S)=\EbpSym(S,R)$.

Note that for \GCb\% as well as for the base pair energy features, the contributions depend on the nucleotides at only a small, constantly bounded number of sequence positions (one and two positions, respectively). This is important for the efficiency of our approach, which is designed to profit from contributions that each depend on only few sequence positions. 
We emphasize that nevertheless, the same framework can capture more expressive RNA energy models like the Turner energy model~\citep{Turner2009}, since such models generally define a function $E$ that maps a pair of a sequence $S$ and a structure $R$ to a real value, the \Def{energy} $E(S,R)$. Here, our approach automatically adapts to the more complex energy functions, leading to a controlled increase of the computational complexity.

Typically, such RNA energy models consider a sequence $S$ \Def{valid} for $R$, iff $\{S_i,S_j\}\in \B$ for
all $(i,j)\in R$, where $\B:=\left\{\{\Ab,\Ub\},\{\Gb,\Cb\},\{\Gb,\Ub\}\right\}$ describes the valid base pairs.

%Typically, such \Def{energy functions} are additively composed from \Def{energy contributions} of structure elements (like single base pairs or loops). It is most instructive, and in this context already practically useful, to develop our approach for the simple \Def{base pair energy model}. This model defines energy contributions $\EbpSym(x,y)$ for pairs $(x,y)\in\Sigma^2$ that specify the energy function as
%$
%\EbpSym(S,R) = \sum_{(i,j)\in R} \EbpSym(S_i,S_j)
%$
%for each sequence $S\in\Sigma^n$ and structure $R$ for length $n$.

%As the structural targets of our design, we fix the set $\R:=\{R_1, \dots, R_k\}$ of $k$ RNA structures for sequences of length $n$.
%Recall that we aim to generate random sequences that satisfy a number of constraints like the compatibility of paired bases. Moreover, we want to enforce complex properties, which we express in terms of \emph{features}. An example of such a feature is the \GCb\% (the corresponding property of a sequence is to have a specific \GCb\%). The additional properties that we are going to enforce are specific 'target' energies for the single structures $R_1, \dots, R_k$; for this purpose, each such energy is considered as a feature.

Finally, to stress the analogy to, in the RNA context more familiar (e.g.~\citep{McCaskill1990}), Boltzmann distributions over the structure ensemble of a sequence $S$, we emphasize the equivalence
\begin{equation}
\label{eq:feature-energy-transformation}
\prod_{1\leq \ell \leq k} \pi_\ell^{-F_\ell(S)} = exp(-\beta E(S)),
\end{equation}
where $\beta$ denotes inverse temperature and $E(S)$ is defined as linear combination of energies $E(S)=\sum_{1\leq \ell \leq k} \ln(\pi_\ell)/\beta E(S,R_\ell)$,
where the energy $E(S,R_\ell)$ is treated as feature $F_\ell(S) = \EbpSym(S,R_\ell);$ as elaborated later, even Turner energies or pseudoknot energies can be expressed as features due to their additive nature.

\paragraph{Dependency structure and complexity}
%The complexity of the central problem depends fundamentally on the dependency structure that is imposed by the entire set of contributions.

%For design in the base pair energy model or the related counting problem, the dependencies are immediately derived from the set of target structures $\R$.

First, for capturing the minimal
dependency structure of multi-target RNA design, which is induced by the requirement of canonical base pairing,
we define the \Def{base pair dependency graph} $G_{\R}$ as the graph with nodes $\{1,\dots,n\}$
and edges $\bigcup_{\ell\in[1,k]} R_\ell$.

More generally, we consider dependencies induced by the complete set of contributions $\F := \bigcup_{\ell}\{f\mid f\text{ contribution of } F_\ell\}$.
%
%Let us denote the \emph{dependencies of a contribution $f$} as \dep{f}.
%
The set $\F$ of contributions induces a
  \Def{dependency graph} $G_\F$ on sequence positions by interpreting the dependencies as hyperedges, namely the hypergraph 
  $G_\F=(\{1,\dots,n\} ,\{\dep(f)\mid f\in \F\})$. The \Def{tree decomposition} of $G_\F$ is fundamental for our algorithms.

\begin{definition}[Tree decomposition and width]
  \label{def:treedecomp}
  Let $G=(X, E)$ be a hypergraph. A \Def{tree decomposition} of $G$ is
  a pair $(T,\chi)$, where $T$ is an unrooted tree/forest, and (for
  each $v\in T$) $\chi(v)\subseteq X$ is a set of vertices assigned to
  the node $v\in T$, such that
\begin{enumerate}
\item each $x\in X$ occurs in at least one $\chi(v)$;
\item for all $x\in X$, $\{ v \mid x \in \chi(v) \}$ induces a connected subtree of $T$;
\item for all $e\in E$, there is a node $v\in T$, such that $e\subseteq\chi(v)$.
\end{enumerate}
The \Def{width} of a tree decomposition $(T,\chi)$ is defined as
$\width(T,\chi) = \min_{u\in T} |\chi(u)| - 1 $. The \Def{treewidth}
of $G$ is the smallest width of any tree decomposition of $G$.
\end{definition}
%Recall the essential role of the treewidth as fixed parameter in our FPT appraoch.

\subsection*{Fixed-parameter tractable (FPT) algorithm}
\label{sec:FPT}

%We express our algorithms in the formalism of constraint networks, here
%specialized as RNA design network. This abstraction allows us to base our
%algorithm on cluster tree elimination (CTE)~\citet{Dechter2013}.

%The RNA design network describes how to evaluate RNA sequences and, in
%particular, partial RNA sequences, which is

To formalize our algorithms, which iteratively merge evaluations of partial solutions, we extend the idea of atomic feature contributions, which are evaluated at sets of the form $\{x_1\mapsto v_1, \dots, x_d\mapsto v_p\}$. Let us call the latter object \Def{partial sequence}; they are handy to specify partial knowledge on the entire sequence at some point of the algorithm. Easily, we can extend the definition of contributions $f$ to sets $\{x_1\mapsto v_1, \dots, x_p\mapsto v_p\}$, where $\{x_1\dots x_p\}$ is any super-set of $\dep(f)$ by ignoring the superfluous assignments $x\mapsto v$, where $x\not\in\dep(f)$.

Our algorithms specialize the idea of cluster tree elimination (CTE)~\citep{Dechter2013}, which operates on constraint networks. In this correspondence, (partial) sequences specialize (partial) assignments and the constraint network would be given by variables for each sequence position, constraints due to valid base pairing, and the set of atomic feature contributions $\cal F$.

%\begin{definition}
%An \Def{RNA design network} (for sequences of length $n$) is a tuple $\network=(\X,\F)$, where
%\begin{itemize}
%\item $\X$ is the set of sequence positions $1,\dots,n$
%\item $\F$ is a set of \Def{contributions} $f:\Sigma^n\to\real$
%\end{itemize}
%\end{definition}
%
%Notably the term `network' refers to the implied graph structure, namely the dependency graph.

%\newcommand{\weighttransform}[2]{{#1}_{||{#2}}}

For algorithmic treatment, it is convenient to encode the weight $\pi$ of each feature in its contributions, such that all contributions can be handled uniformly. This transformation works by multiplying all contributions with $ln(\pi)$, where $\pi$ is the weight of the corresponding feature, since then
$exp(-\ln(\pi)f(S))=\pi^{-f(S)}.$

%For this purpose, we define the \Def{transformed contribution} 
%$\weighttransform{f}{\pi}(S) := \ln(\pi)f(S),$
%such that $exp(-\weighttransform{f}{\pi}(S))=\pi^{-f(S)}.$ 
%
Let us now specify the concrete set $\F$ of contributions that we use
for the design in the base pair model targeting \GCb\% and structures $\R$ with weights $\pi_0,\dots,\pi_k$. The set $\F$ thus consists of
\begin{itemize}
\item the transformed contributions $\ln(\pi_0)f^{\GCb}_i$ for the ~$\GCb\%$ feature ($i=1,\dots,n$)
\item the transformed contributions $\ln(\pi_\ell)f^{R_\ell}_{ij}$for each structure $R_\ell\in\R$ and $(i,j)\in R_\ell.$
\end{itemize}

By these definitions, the set $\F$ encodes the partition function $\partfun{\pi_0,\dots,\pi_k}$ of Eq.~(\ref{eq:mainproblem}); also confer Eq.~(\ref{eq:feature-energy-transformation}). 
%Subsequently, we show how to efficiently compute this partition function based on the 
%corresponding dependency graph and its tree decomposition.

\subsubsection*{Partition function and stochastic backtracking}\label{sec:PF}

We compute the partition function (as specified by $\F$) by dynamic programming based
on a tree decomposition of $\F$'s dependency graph.
Note that analogous algorithms could be constructed to count valid sequences or minimize energy.
%This yields fixed parameter tractable (FPT) algorithms, based on the treewidth.

%\begin{figure}[t]
%{\centering\scalebox{11}{\fbox{$\phantom{5689}$}}\\}
%
%  \caption{Figure illustrating RNA network to hypergraph to tree decomposition}
%\end{figure}

Our algorithms are formulated to process a \Def{cluster tree} of $\F$, which is a tuple $(T,\chi,\phi)$, where
$(T,\chi)$ is a tree decomposition of $G_\F$, and $\phi(v)$ represents
a set of functions $f$, each uniquely assigned to a node $v\in T$;
$\dep(f)\subseteq\chi(v)$ and $\phi(v)\cap \phi(v')=\varnothing$ for
all $v\neq v'$.

Two further notions are essential for our algorithms: for two nodes $v$ and $u$ of the cluster tree, define
their \Def{separator} as $\separator{u}{v} := \chi(u)\cap\chi(v)$;
moreover, we define the \Def{difference positions} from $u$ to an
adjacent $v$ by $\difference{u}{v}:=\chi(v) - \separator{u}{v}$.

Since our algorithms iterate over specific sets of sequence positions, we moreover define
the \Def{set $\partseqs(\Y)$ of all partial sequences determining the positions of $\Y=\subseteq\{1,\dots,n\}$} in all combinations of nucleotides $\{\Ab,\Cb,\Gb,\Ub\}$,
i.e.~for $\Y=\{y_1,\dots,y_{q}\}$,
\begin{align*}
\partseqs(\Y) = \{ & \{ y_i\mapsto v_i \mid i=1,\dots,q \} \\& | (v_1,\dots,v_{q}) \in \{\Ab,\Cb,\Gb,\Ub\}^{q} \}.
\end{align*}
%
We assume several properties of the given cluster tree (reflecting $\F$), which otherwise could be easily obtained:
\begin{itemize}
\item
$T$ is
connected and contains a dedicated node $r$, with
$\chi(r)=\varnothing$ and $\phi(r)=\varnothing$. If such a root does not exist, it can be added to the tree decomposition and connected to one node in each connected component of $T$.
\item
all edges in the tree decompostion are oriented towards this root.
\item
all sets $\difference{u}{v}$ are singleton: for any given
cluster tree, an equivalent (in term of treewidth) cluster tree can
always be obtained by inserting at most $\Theta(|\X|)$ additional
clusters.
\end{itemize}

Algorithm~\ref{alg:pf} computes the partition function by passing
messages along the directed edges $u\to v$ (which point from 
child $u$ to its parent $v$). Each message $m$ has the form of a contribution, i.e. it takes a partial sequence, depends
on the positions $\dep(m)\subseteq \X$, and yields a partition function
in $\real$. The message from $u$ to $v$ represents the
partition functions of the subtree of $u$ for all possible partial
sequences in $\partseqs(\separator{u}{v})$. Induction over $T$ lets us show
the correctness of the algorithm
(Supp. Mat.~\ref{appsec:correctness}).  After running
Alg.~\ref{alg:pf}, multiplying the 0-ary messages sent to the root $r$
yields the total partition function (i.e.~due to proper encoding the partition function of our design problem)
\begin{math}
  \prod_{(u\to{}r)\in T} \evalfor{\Message{u}{r}}{\varnothing}.
\end{math}

\begin{algorithm}[t]
  \KwData{Cluster tree $(T,\chi,\phi)$} \KwResult{Messages
    $\Message{u}{v}$ for all $(u\to{}v)\in T$; i.e.~partition
    functions of the subtrees of all $v$ for all possible partial
    sequences determining exactly the positions $\separator{u}{v}$.}
  \For{$u\to{}v\in T$ in postorder}{
    \For{$\val\in\partseqs(\separator{u}{v})$}{
      $x\gets 0$\;
      \For{$\val'\in\partseqs(\difference{u}{v})$}{
        $p \gets$ product( $exp(-\evalfor{f}{\substitute{\val}{\val'}})$ for $f\in \phi(u)$ )\\
        ${}\quad\qquad \cdot\ $product( $\evalfor{\Message{u'}{u}}{\substitute{\val}{\val'}}$ for $(u'\to{}u)\in T$ )\;
        $x \gets x + p$\;
      }
      $\evalfor{\Message{u}{v}}{\val} \gets x$\;
    }
    \Return {$m$}\;
  }
  \caption{FPT computation of the partition function using dynamic
    programming, i.e.~cluster tree elimination (CTE). The postorder traversal guarantees that
    when processing edge $u\to{}v$, all messages
    $\Message{u'}{u}$, corresponding to DP matrices, have been computed before. }\label{alg:pf}
\end{algorithm}

The partition functions can then direct a stochastic backtracking procedure to sample sequences from the Boltzmann distribution (according to $\F$).
%
For an expanded cluster tree, after the messages $\Message{u}{v}$ for
the edges in the tree decomposition are generated by Algorithm \ref{alg:pf},
one can repeatedly call Algorithm~\ref{alg:sampling}, each time
randomly drawing another sequence from the Boltzmann distribution.

\subsection*{Computational complexity of the multiple target sampling algorithm}\label{sec:complexity}

%The complexity of the proposed sampling algorithm depends on the treewidth of the dependency graph $G_\F$.
% The treewidth $w$ of a cluster tree $(T,\chi,\phi)$, $T=(V,E)$, is $\max_{v\in V} |\chi(v)| - 1$, i.e.~the maximum number of positions in any of its clusters minus 1.
Let $s$ denote the maximum size of any separator set $\separator{u}{v}$ and $D$ denote the maximum size of
$\difference{u}{v}$ over $(u,v)\in E$.  In the absence
of specific optimizations, running Alg.~\ref{alg:pf} requires
$\mathcal{O}((|\F|+|V|)\cdot 4^{w+1})$ time and
$\mathcal{O}(|V|\cdot4^s)$ space
(Supp. Mat.~\ref{appsec:algcomplexity}); Alg.~\ref{alg:sampling} would
require $\mathcal{O}((|\F|+|V|)\cdot 4^D)$ per sample on arbitrary
tree decompositions
(Supp. Mat.~\ref{appsec:algcomplexity}). W.l.o.g.\ we assume that
$D=1$; note that tree decompositions can generally be transformed,
such that $\difference{u}{v}\leq 1$.
%
Moreover, the size of $\F$ is linearly bounded: for $k$
input structures for sequences of length $n$, the energy function is
expressed by $\mathcal{O}(n\,k)$ functions. Finally, the number of cluster
tree nodes is in $O(n)$, such that $|\F|+|V| \in \mathcal{O}(n\,k)$.

% \begin{algorithm}
%  \KwData{Node $u$, partial sequence $\val\in\partseqs(\separator{u}{v})$;\newline
%  Cluster tree $(T,\chi,\phi)$ and partition functions $\Message{u'}{v'}[\val']$, $\forall (u'\to{}v')\in T$ and $\val'\in\partseqs(\separator{u'}{v'})$.}
%  \KwResult{Boltzmann-distributed random partial sequence for the subtree rooted at $u$, specializing a partial sequence $\val$.}
%  \Fn{\Sample$(u,\val)$}{
%    $r \gets \Random(\evalfor{\Message{u}{v}}{\val})$\;
%    \For{$\val'\in\partseqs(\difference{u}{v})$}{

%      $p \gets$ product( $exp(-\evalfor{f}{\substitute{\val}{\val'}})$ for $f\in \phi(u)$ )\\
%      ${}\quad\qquad \cdot\ $product( $\evalfor{\Message{u'}{u}}{\substitute{\val}{\val'}}$ for $(u'\to{}u)\in T$ )\;
%   	  $r \gets r - p$\;
%   	  \If{$r<0$}{
%   	    $\val_{\rm res} \gets \substitute{\val}{\val'}$\;
%   	  \For{$(u'\to{}u)\in T$}{
%   	      $\val_{\rm res} \gets \substitute{\val_{\rm res}}{\Sample(u',\substitute{\val}{\val'})}$\;
%      }
%   	  \Return {$\val_{\rm res}$}
%   	  \;}
%  }
%  }
%  \caption{Stochastic backtrack algorithm for partial sequences in the Boltzmann distribution.}\label{alg:sampling}
% \end{algorithm}

\begin{algorithm}
 \KwData{Cluster tree $(T,\chi,\phi)$ and partition functions $\Message{u'}{v'}$ for all $(u'\to{}v')\in T$.}
 \KwResult{Boltzmann-distributed random sequence $\val$.}
 $\val$ \gets $\varnothing$\;
 \For{$u\to{}v\in T$ in preorder}{
   $r \gets \text{uniform random number between $0$ and $\evalfor{\Message{u}{v}}{\val}$}$\;
   \For{$\val'\in\partseqs(\difference{u}{v})$}{

     $p \gets$ product( $exp(-\evalfor{f}{\substitute{\val}{\val'}})$ for $f\in \phi(u)$ )\\
     ${}\quad\qquad \cdot\ $product( $\evalfor{\Message{u'}{u}}{\substitute{\val}{\val'}}$ for $(u'\to{}u)\in T$ )\;
     $r \gets r - p$\;
     \If{$r<0$}{
       $\val \gets \substitute{\val}{\val'}$\;
     }
   }
 }
 \Return {$\val$}\;

 \caption{Stochastic backtrack algorithm for partial sequences in the
   Boltzmann distribution. Processing the edges $u\to{}v\in T$ in
   preorder ensures that $\val$ invariantly determines all positions
   of $v$ outside the subtree of $u$.}\label{alg:sampling}
\end{algorithm}

%For example, for design to a single non-crossing target in the nearest neighbor energy model, the complexity of Algs.~\ref{alg:pf} and \ref{alg:sampling} degrades to linear time (as reported for IncaRNAtion), since there are only $O(n)$ many functions and nodes (in the sequence length $n$). Moreover, the dependencies are tree-like owing to the tree-like non-crossing structure; this implies a constantly bounded maximum treewidth.

\begin{theorem}[Complexities]\label{th:complexities}
  For sequence length $n$, $k$ target structures and treewidth $w$,
  % and a base pair dependency graph having $c$ connected components%
  $t$ sequences
  are generated from the Boltzmann distribution  in
  $O( n\, k \, 4^{w+1} + t\, n\, k )$ time.
  %$O( n\, k \min(2^{w+c+1},4^{w+1}) + t\, n\, k )$ time.
  %$\mathcal{O}( 2^d\, n\, k  + t\, n\, k )$ time, where $d:=\min(w+c+1,2(w+1))$.
\end{theorem}

By this theorem, Boltzmann sampling in our setting is fixed parameter tractable (FPT) in the tree width $w$ (since the complexity is polynomial for fixed $w$).
%
As shown in Supp. Mat.~\ref{sec:improvedComplexity}, the complexity of the precomputation can be further improved to
$\mathcal{O}(n\,k\,2^{w+1}\,2^{c})$, where $c$ is the maximum number of connected components represented in a node of the tree decomposition ($c\le w+1$).

Note that in this complexity analysis, we do not include time and space for computing the
tree decomposition itself, since we observed that the computation time
of tree decomposition (\Software{GreedyFillIn}, implemented in
\Software{LibTW}~by \citet{Dijk2006}) for multi-target sampling is
negligible compared to Alg.~\ref{alg:pf}
(Supp. Mat.~\ref{appsec:treedecomp} and
\ref{appsec:dependency-cliques}).

\subsection*{Design for expressive energy models}\label{sec:energy_models}

% \paragraph{Nearest-neighbor model.}
% Let us now consider a simplified version of the nearest-neighbor model, where the free energy contribution  $\Ehp{x,y,s}$ of a hairpin only depends on the closing base pair $(x,y)$, and the contribution $\Eint{x,y,x',y',s}$ of an interior loop depends on its opening/closing bases pairs $(x,y)/(x',y)$, and its length $s$. Moreover, the energy of a multi loops with $s$ inner base pairs and $t$ unpaired base pairs is approximated as $a+b\,s+c\,t$, where $a,b,c$ are predefined constants.

% Again, the energy is expressed by adding functions to $F$ for each $R_\ell$ and base pair $(i,j)\in R_\ell$:
% \begin{itemize}
%   \item if $(i,j)$ closes a hairpin in $R_\ell$, add $f$, s.t.~$\evalfor{f}{\val_S}=\Ehp{S_i,S_j,j-i}$
%   \item if $(i,j)$ closes an interior loop, bulge or stack with inner base pair $(i_1,j_1)$ in $R_\ell$, add $f$, s.t.~$\evalfor{f}{\val_S}=\Ehp{S_i,S_j,S_k,S_l,i_1-i+j_1-j}$
%   \item if $(i,j)$ closes a multiloop add $f$, s.t.~$\evalfor{f}{\val_S}=a$.
%   \item if $(i,j)$ is inner base pair of a multiloop of $R_\ell$ add $f$, s.t.~$\evalfor{f}{\val_S}=b$.
% \end{itemize}
% Finally add for each $R_\ell$ and unpaired base $i$ in a multi loop of $R_\ell$, the function $f$, $\evalfor{f}{\val_S}=c$.

So far, we discussed sampling only in base pair energy models, although one is finally interested in targeting energies in realistic energy models like the Turner model or pseudoknot models like {\tt HotKnots}~\citep{Ren2005}.
Notably, for this purpose, the discussed sampling strategy can be extended in two ways: 1) one can directly sample based on more expressive energy models, 2) sampling in a simple energy model can be used to approximate sampling in more complex models.

\paragraph{Sampling based on more expressive energy features}
Complex \Def{loop-based}
 energy models---e.g.~the Turner model,
which among others includes energy terms for special loops and dangling
ends---can be encoded as straightforward extension of our outline of the base pair energy model.

In practice, the energy model has a strong influence on the computational complexity of our approach. Therefore, it is interesting to consider, in addition to the base pair model, other
stripped-down variants of the nearest neighbor model, which could offer a compromise between the low-complexity base pair energy model and the high-accuracy Turner model. 
Such a model, which is particularly promising for this purpose, is the
\Def{stacking energy model}. This model assigns non-zero energy
contributions only to stacks, i.e. interior loops with closing base
pair $(i,j)$ and inner base pair $(i+1,j-1)$.

The arity of the introduced functions bounds the
treewidth of the dependency graph. %(and therefore computational complexity).
Thus, it is noteworthy that the base pair energy model
requires only binary functions; the stacking model, only quarternary
dependencies. This arity is increased in a few cases by the commonly
used Turner 2004 model~\citep{Turner2009} for encoding tabulated
special hairpin and interior loop contributions, which depend on up to
nine bases for the interior loops with a total of 5 unpaired bases
(``2x3'' interior loops)---all other energy contributions (like
dangling ends) still depend on at most four bases of the sequence.

\paragraph{Approximating Turner Energy using Simpler Energy Models}
To capture the realistic Turner model $\EnergyTurner$ more efficiently, we exploit the tight correlation between
$\EnergyTurner$ and the fitted stacking model $\EnergyStacking$
(Supp. Fig~\ref{fig:training-cor}). More precisely, we observed a
structure-specific affine dependency between the Turner and stacking
energy models, so that
$\EnergyTurner(S;R) \approx \gamma\cdot \EnergyStacking(S;R) + \delta$
for any structure $R$ and sequence $S$. We inferred the
$(\gamma,\delta)$ parameters from a set of sequences generated with
homogenous weights $w=e^{\beta}$, tuning only \GCb\% to a
predetermined value.  Finally, we adjusted the targeted energies
within our stacking model to
$\EnergyStacking^{\star} = (\EnergyTurner^{\star}- \delta)/\gamma$ in
order to reach, on average, the targeted energy
$\EnergyTurner^{\star}$ in the Turner model.

\subsection*{Extension to multidimensional Boltzmann sampling}\label{sec:multiBoltzmann}
The flexibility of our framework supports the advanced sampling technique "multidimensional Boltzmann sampling"~\citep{Bodini2010}, which allows (probabilistically) enforcing additional, complex properties of the samples.
This technique was previously used to control \GCb\%~\citep{Waldispuehl2011,Reinharz2013} and dinucleotide content~\citep{Zhang2013} of sampled RNA sequences; here, in addition to controlling \GCb\% (our feature $F_0$) we use it to target the free energies $(\TargetE_1,\ldots,\TargetE_k)$ of the single targets (features $F_1,\dots,F_k$).

For the multidimensional Boltzmann sampling, we require the already established ability to \Def{sample from a weighted distribution} over the set of valid sequences, where the probability of a sequence $S$ is
$\mathbb{P}(S\mid \pmb{\pi}) = \frac{\prod_{\ell=0}^{k} \pi_i^{-F_i(S)}}{\partfun{\pmb{\pi}}},$
where $\pmb{\pi}:=(\pi_0\cdots\pi_k)$ is the vector of the positive real-valued \Def{weights}, and $\partfun{\pmb{\pi}}$ is the weighted partition function.

%Such a distribution can be induced by a simple modification of the functions described in Sec.~\ref{sec:energy_models}, where any energy function $E(\val)$ for a structure $\ell$ is replaced by $E'(\val):= \ln(\pi_\ell)\, E(\val)/\beta.$ 
Due to the construction of $\F$,
the probability of a sequence $S$ is thus proportional to
$\prod_{\ell=0}^{k} \pi_i^{-F_i(S)} = \prod_{f\in\F} f(S).$

One then needs to \Def{learn a weights vector} $\pmb{\pi}$ such that, on average, the targeted energies are achieved by a random sequences in the weighted distribution, i.e.\ such that  $\mathbb{E}(F_\ell(S)\mid \pmb{\pi})=\TargetE_\ell$,  $\forall\ell\in[1,k]$ and, analogously, the expectation of $F_0(S)$ is the targeted GC content.
The expected value of $F_\ell$ is always decreasing for increasing weights $\pi_\ell$ (see Supp. Mat.~\ref{sec:weight_derivatives}). More generally, computing a suitable parameter vector $\pmb{\pi}$ can be restated as a convex optimization problem, and be efficiently solved using a wide array of methods~\citep{Denise2010,Bendkowski2017}.
In practice, we use a simple heuristics which starts from an initial weight vector $\pmb{\pi}^{[0]}:=(e^\beta,\dots,e^\beta)$ for $\beta=1/(RT)$, T=$37^\circ$, and gas constant $R$. Then, at each iteration, it generates samples $\mathcal{S}$ of sequences. The expected value of an energy $F_\ell$ is estimated as $\hat\mu_\ell(\mathcal{S}) = \sum_{S\in\mathcal{S}}F_\ell(S)/|\mathcal{S}|$, and the weights are updated at the $t$-th iteration by %  $\pi_\ell^{[t+1]} = \pi_\ell^{[t]}\cdot\TargetE_\ell/\hat\mu_\ell(\mathcal{S})$.
$\pi_\ell^{[t+1]} = \pi_\ell^{[t]}\cdot \gamma^{\hat\mu_\ell(\mathcal{S})-\TargetE_\ell}$. In practice, the constant $\gamma>1$ is chosen empirically  ($\gamma=1.2$) to achieve effective optimization.
While heuristic in nature, this basic iteration was elected in our initial version of \ourprog{} because of its good empirical behavior.

A further \Def{rejection step} is applied to retain only those sequences whose energy for each structure $R_\ell$ belongs to $[\TargetE_\ell\cdot(1-\varepsilon),\TargetE_\ell\cdot(1+\varepsilon)]$, for $\varepsilon\ge 0$ some predefined \Def{tolerance}. The rejection approach is justified by the following considerations:
i) \emph{Enacting an exact control over the energies would  be technically hard and costly.} Indeed, controlling the energies through dynamic programming would require explicit convolution products, generalizing~\citep{Cupal1996}, inducing additional $\Theta(n^{2k})$ time and $\Theta(n^k)$ space overheads;
ii) \emph{Induced distributions are typically concentrated.} Intuitively, unless sequences are fully constrained individual energy terms are independent enough so that their sum is concentrated around its mean -- the targeted energy (cf.{} Fig.~\ref{fig:energydist-pk}).
%Indeed, as shown in Fig.~\ref{fig:shifting_mean}, empirical joint distributions in the energies have good fits towards Normal multidimensional distributions with linear (co-)variances.
For base pair-based energy models and special base pair dependency graphs
(paths, cycles\ldots) this property rigorously follows from analytic
combinatorics, see \citep{Bender1983} and
\citep{Drmota1997}. In such cases, the expected number of
rejections before reaching the targeted energies remains constant when
$\varepsilon\ge 1/\sqrt{n}$, and $\Theta(n^{k/2})$ when
$\varepsilon=0$. 
%The \GCb\% of designs can also be controlled,
%jointly and in a similar way, as done in
%\Software{IncaRNAtion}~\citep{Reinharz2013}.

\subsection*{\#{\sf P}-hardness of counting valid designs}\label{sec:counting}
%%% reformulate later
%We define the following notations: $\real^\infty:=\real\cup\{\infty\}$, $[i,j]:=\{i,\ldots,j\},$ \ldots
Finally, we turn to the complexity of $\NumDesign(G)$, the problem of
computing the number of valid sequences for a given compatibility
graph $G=(V,E)$. Note that this problem corresponds to the partition
function problem with weights $\pi_\ell=1$ for each target structure, so that each valid structure contributes $1$ to the partition
function; each invalid structure, 0.

As previously noted~\citep{Flamm2001}, a set of target
structures admits a valid design iff its compatibility graph is
bipartite, which can be tested in linear time.  Moreover, without
$\{\Gb, \Ub\}$ base pairs, the design of any connected component $C$
of $G$ would be entirely determined by the assignment of a single of
its nucleotides. The number of valid designs is thus simply
$4^{\#{\rm CC}(G)}$, where $\#{\rm CC}(G)$ is the \Def{number of
  connected components}.
The problem becomes much harder when $\{\Gb, \Ub\}$ base pairs are allowed, and we show that valid designs for a set of structures cannot be counted in polynomial time, unless ${\sf\# P}={\sf FP}$. This equality would, in particular, imply the classic ${\sf P}={\sf NP}$.

To establish that claim, we consider instances $G=(V_1\cup V_2, E)$ that are connected and bipartite ($E \cap (V_1\times V_2) = E$), noting that hardness over restricted instances implies hardness of the general problem. Moreover note that, as we explained before, assigning a nucleotide to a position $u\in V$ constrains the parity ($\{\Ab,\Gb\}$ or $\{\Cb,\Ub\}$) of all positions in the connected component of $u$. For this reason, we restrict our attention to the counting of valid designs \emph{up to trivial  symmetry} $(\Ab\leftrightarrow \Cb/\Gb\leftrightarrow \Ub)$, by constraining the positions in $V_{1}$ to $\Ab$ and $\Gb$. Let $\Design{G}$ denote the subset of all designs for $G$ under this constraint, noting that $\NumDesign(G) = 2\cdot|\Design{G}|$.

Let $\IS{G}$ denote the set of all independent sets in the connected graph $G$; recall that an \Def{independent set} of $G=(V,E)$ is a subset $V'\subseteq V$ of nodes that are not connected by any edge in $E$.

\begin{proposition}\label{prop:bijection}
  $|\Design{G}| = |\IS{G}|$.
\end{proposition}
%
\begin{proof}
Consider the mapping $\Psi: \Design{G} \to \IS{G}, f \mapsto \left\{v\in V\mid f(v)\in\{\Ab,\Cb\}\right\}.$
We show that $\Psi$ is bijective:%
\begin{itemize}
\item $\Psi$ is injective, i.e. $\Psi(f)\neq\Psi(f')$ for all $f\neq f'$.
If $f\neq f'$, then there exists a node $v\in V$ such that $f(v)\neq f'(v)$.
We discuss only the case $v\in V_1$, where we restricted the nucleotides to $\Ab$ and $\Gb$. Then, $\{f(v),f'(v)\}$ must equal $\{\Ab,\Gb\}$, such that either $v\in\Psi(f)$ or $v\in\Psi(f')$.
%
b\item $\Psi$ is surjective, i.e. there is a preimage for each element $I\in \IS{G}$. Define $f\in\Design{G}$ as\begin{displaymath}
f(v) = \left\{\begin{array}{ll} % @{\qquad}ll
\Ab{} & \text{if $v\in V_1$ and $v\in I$} \\ %&
\Cb{} & \text{if $v\in V_2$ and $v\in I$} \\
\Gb{} & \text{if $v\in V_1$ and $v\not\in I$} \\ %&
\Ub{} & \text{if $v\in V_2$ and $v\not\in I$}
\end{array}\right.
\end{displaymath}
One easily verifies that $\Psi(f) = I$.
%
%for any $(v_1,v_2)\in V_1\times V_2$ by
%\begin{align*}
%f(v_1) &= \begin{cases} \Ab & \text{if }v_1\in S\\ \Gb & \text{if }v_1\notin S\end{cases} &
%\text{\quad and\quad}&&
% f(v_2) &= \begin{cases} \Cb & \text{if }v_2\in S\\ \Ub & \text{if }v_2\notin S.\end{cases}
%\end{align*}
%
It remains to show that $f$ is a valid design for $G$, i.e. for each $(v, v') \in E$, $\{f(v),f(v')\}\in \B$; please recall that we defined $\B$ as the set of all valid nucleotide pairs.
Assume there is an edge $(v_1,v_2)\in E$, violating $\{f(v_1),f(v_2)\} \in \B$. Since $G$ is bipartite,
$v_1\in V_1$ and $v_2\in V_2$, such that $f(v_1)\in\{\Ab,\Gb\}$ and $f(v_2)\in\{\Cb,\Ub\}$. This implies that among all possible $\{f(v_1), f(v_2)\}$ only $\{\Ab,\Cb\}$ is not in $\B$, which in turn requires $v_1\in I$ and $v_2\in I$. Therefore, since $I$ is an independent set, the edge $(v_1,v_2)\in E$ cannot exist. 
\end{itemize}
\end{proof}

Now we can build on the connection between the two problems to obtain complexity results for \NumDesign. Counting independent sets in bipartite graphs ($\#{\sf BIS}$) is indeed a well-studied \#{\sf P}-hard problem~\citep{Ge2012}. Now assume the existence of an efficient (polynomial-time) algorithm $\mathcal{A}$ for computing $|\NumDesign(G)|$ on connected (bipartite) graphs. Then, running $\mathcal{A}$ and returning $|\NumDesign(G)|/2$ constitutes an efficient algorithm for $\#{\sf BIS}$ on connected graphs.
However, $\#{\sf BIS}$ is also \#{\sf P}-hard on connected graphs, as the number of independent sets for a disconnected graph $G$ can be efficiently computed by $|\IS{G}|=\prod_{cc\in CC(G)} |\IS{cc}|$. In other words, any efficient algorithm for \NumDesign implies an efficient algorithm for $\#{\sf BIS}$, and we conclude with the hardness of \NumDesign.

\begin{corollary}
  \NumDesign is $\#{\sf P}$-hard
\end{corollary}
% \begin{proof}

%   Given any algorithm $\mathcal{A}$ that solves \NumDesign over strongly-connected graphs $G$, one can  construct an algorithm $\mathcal{A}'$ that solves $\#{\sf BIS}$:
%   \begin{enumerate}
%     \item[1)] calculate $\NumDesign(G)$ by calling $\mathcal{A}$
%     \item[2)] return\\$\NumDesign(G)/2=|\Design{G}| = |\IS{G}|$.
%   \end{enumerate}
%     Obviously, if $\mathcal{A}$ is a polynomial-time algorithm, so is $\mathcal{A'}$.
%   This implies that $\NumDesign$  is at least as hard as $\#{\sf BIS}$, thus does not admit a polynomial time exact algorithm unless $\#{\sf P}={\sf FP}$.
% \end{proof}

\section*{Results}\label{sec:results}
Our Boltzmann sampling strategy (Algorithms
\ref{alg:pf} and \ref{alg:sampling}) was implemented, using a combination of {\tt C++} and {\tt Python}, in a software called \ourprog{}, available at:

{\centering \url{https://github.com/yannponty/RNARedPrint}\\[.3em]}

The driver script of \ourprog{} takes as input a set of (secondary) structures, ... and outputs ....

We implemented the core algorithms in C++, resulting in the tool
C++-tool \ourprog{}, which takes a set of target structures, as well as
weights for each energy feature and \GCb\%; then,
it generates a sample (of prescribed size) from the corresponding
Boltzmann distribution. For \ourprog{}, we implemented the base pair
energy model and the stacking energy model.
Moreover, we provide a Python wrapper script for the C++-tool
\ourprog{} to perform targeting, based on multi-dimensional Boltzmann
sampling.
For given secondary structures, together with prescribed target energies
and target \GCb\%, this script generates
a series of sequences that satisfy the target values for the energy and
\GCb\% features within configurable tolerances.
Notably, these target energies are actual free energies in the realistic
Turner energy model, which are targeted by efficiently sampling in the
base pair energy model.

\subsection*{Effectively targeting Turner energies}
We implemented our Boltzmann sampling strategy (Algorithms
\ref{alg:pf} and \ref{alg:sampling}), to sample valid sequences for
given target structures and weights $\pi_1,\dots,\pi_k$.  Moreover, we
used multi-dimensional Boltzmann sampling to target specific energies and
\GCb\%.  Our tool \ourprog{} evaluates energies according to the
stacking energy model $\EnergyStacking$, whose parameters were fitted
to best approximate Turner energies. As well, we implemented and
fitted a base pair energy model for \ourprog{}, which was not studied
for its targeting performance (both models:
Supp. Mat.~\ref{appsec:modelparameters}).
%
%\paragraph{Implementation of multi-dimensional Boltzmann sampling}
%As suggested in Section~\ref{sec:multiBoltzmann}, we heuristically optimize weights to target specific energies and \GCb\% in an iterative procedure (based on the core sampling algorithm).
%Specifically, energy weights are initialized as $w_\ell=e^{\frac{1}{RT}}$; the gc weight as $w_{gc}=1$. For each iteration, $\psi\cdot n$ sequences are sampled with the current weights. Then, we determine the mean energies $E(S;R_\ell)$ and the mean GC content over the sampled sequences. For the next iteration, the weights are updated by the rules $w_\ell = w_\ell \gamma^{mean(E(S;R_\ell))-E_\ell}$ and $w_{gc} = w_{gc}\frac{GC_{target}}{mean(GC)}$. During this procedure, we collect   in a set $P$ all the sequences exhibiting energies within $[E_1-\delta,E_1+\delta]\times [E_2-\delta,E_2+\delta]\ldots$ and and GC-content within $[GC_{target}-\idelta',GC_{target}+\delta']$.
%If $|P| > n$ or an maximum amount of iterations is reached, the sequences in $P$ are returned.
%

Fig.~\ref{fig:energydist-pk} illustrates how well complex realistic energy models can be approximated based on simpler, but better tractable ones. For
the two target structures of Fig.~\ref{fig:energydist-pk}B,
Fig.~\ref{fig:energydist-pk}A shows the good fit between realistic
energies in the full-fledged Dirks and Pierce energy model for
pseudoknots (D\&P model) and energies in the stacking energy
model. For the shown fits we sampled $n=1\,000$ sequences, targeting a
\GCb\% of $60\%$.
%
For an example instance of the \texttt{Modena} benchmark with two
pseudoknotted target structures, Fig.~\ref{fig:energydist-pk}B shows
the Turner energy distributions of the single structures as they
result from sampling with different weight parameters. The figure
illustrates how our multidimensionnal Boltzmann sampling strategy can,
to a large extent, independently shift the Turner energies of sampled
sequences towards prescribed targets. See Supp. Mat.,
Fig.~\ref{fig:energydist} for a further example with three
pseudoknot-free target structures.

%To reduce the cost of the precomputation, we used a stacking energy model
%as a proxy for the Turner model. Indeed, we observed a strong linear
%dependency between the Turner energy and the
%emphasize that only the simple stacking energies are directly targeted
%(at $\pm$10\% tolerance, except for a few hard instances). Due to our
%linear adjustment, we finally achieve well-controlled energies in the
%realistic Turner energy model.
\begin{figure*}[t]
  \begin{center}
    %{\sf \bfseries A}\
    \includegraphicstop[width=0.97\textwidth]{Figs/energy_shift}\hfill
    %{\sf \bfseries B}\ \includegraphicstop[width=0.55\textwidth]{Figs/PKB00211_PKB00239_0_energy_distribution}
  \end{center}
  \caption{%
    Targeting specific energies for pseudoknotted structures using
    multi-dimensional Boltzmann sampling. \textbf{(A)} Linear fits between the
    energies in the stacking model to the realistic pseudoknot energy
    model by Dirks and Pierce (D\&P) for sampled sequences; the good
    match enables more efficient targeting of Turner energies based on
    targeting stacking model energies.  We show the fits for both
    target structures R1 and R2 of subfigure B.  \textbf{(B)} D\&P energy
    distributions for two target structures R1 and R2. We show the
    results of targeting the respective free energies $(-30,-20)$,
    $(-30,-30)$, $(-25,-25)$, and $(-20,-30)$ for the target
    structures---this demonstrates the effectivity of our adaptive
    multi-dimensional Boltzmann sampling procedure; moreover, for
    comparison, the distributions for uniform and Boltzmann
    samples---respectively associated with homogenous weights $1$ and
    $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{fig:energydist-pk}
\end{figure*}

\subsection*{Generating high-quality seeds for further optimization}
We empirically evaluated \ourprog{} for generating seed sequences targeting multiple (pseudoknotted) structures, possibly followed by subsequent local optimizations. As a baseline for comparison, we considered \RNAblueprint~\citep{Hammer2017}, the current leading tool for multiple design.
As a quality measure, we applied the objective function that was introduced by~\citet{Hammer2017} for multi-stable design, defined as:
% \begin{align}
%   f(S)= \quad & \frac{1}{k} \sum_{\ell=1}^{k} (E(S, R_\ell) - G(S))\notag\\
%    & +\ \xi \frac{2}{k(k-1)} \sum\limits_{1\leq\ell<j\leq k}|E(S,R_\ell) - E(S,R_j)|.
% \end{align}
\begin{multline}
  \label{eq:blueprintobjective}
    \Obj(S) = \frac{1}{k} \sum_{\ell=1}^{k} (E(S, R_\ell) - \EnsE(S))\\
    +\frac{1}{2\binom{k}{2}} \sum\limits_{1\leq\ell<j\leq k}|E(S,R_\ell) - E(S,R_j)|,
\end{multline}
%
where the free energies $E(S, R)$ as well as the \Def{ensemble free energy} $\EnsE(S)$ of $S$ are computed
by {\tt RNAfold}~\citep{Lorenz2011} in the pseudoknot-free case; for
pseudoknotted targets, $\EnsE(S)$ is approximated by the \Def{minimum
  free energy} (MFE) of $S$ as estimated by {\tt HotKnots}~\citep{Ren2005} in the energy model
of~\citet{dirks-pierce-03}.  Intuitively, the first term of \Obj{}
captures the distance of the targets from the ensemble free energy,
while the second term penalizes the dispersion of targets; \Obj{} is
best (minimal) when all targets simultaneously achieve the minimum
free energy of the sequence.

We considered a benchmark of six sets of target structures described in \citep{Taneda2015}: {\sf 2str}, {\sf 3str}, and {\sf 4str} consist of non-pseudoknotted structures, while {\sf PK60}, {\sf PK80}, and {\sf LE80} contain pseudoknotted structures.
%%For each instance of the \Software{Modena}
%benchmark~\citep{Taneda2015}, we generated $1\,000$ seed sequences at these energies.
%Our procedure is tailored to produce sequences with similar Turner
%energy that favor the stability of the target structures having
%moderate \GCb\%; all these properties are desirable objectives
%for RNA design.
Using \ourprog{}, we generated at least 1\,000 seeds for
each instance of the benchmark. We used target energies that were
computed by averaging the energy over samples at higher weights
$e^{\beta}$, and set the targeted \GCb\% to 60\%.  We compared the
\Obj{} value of these sequences against that of seed sequences,
uniformly sampled using \RNAblueprint{}.  Moreover, for both sets, we
used an adaptive greedy walk~\citep{Hammer2017} to \Def{minimize} the
\Obj{} function. At each step, the local search resamples (uniformly
at random) the positions of a randomly selected component in the base
pair dependency graph, accepting the modification only if it results
in a gain. We performed 500 greedy descent steps in the case of
pseudoknot-free datasets {\sf 2str}, {\sf 3str}, and {\sf 4str}; and
200 steps for the pseudoknotted ones {\sf PK60}, {\sf PK80}, and {\sf
  LE80}.

The results, shown in Fig.~\ref{fig:benchmark-results}, reveal that
Boltzmann-sampled sequences outperform uniform seeds on every dataset,
leading to average improvements in \Obj{} values ranging from $7.26$
({\sf LE80}) to $16.05$ ({\sf 2str}) units. Remarkably, this
improvement is be observed for both of the terms in \Obj{} (see
Supp. Mat.~\ref{sec:improvements}).  This means that \ourprog{}
produces sequences whose targets are both substantially closer to
the sequence ensemble/minimal free energy, and more similarly stable across targets.  In fact, for
every sequence in our benchmark, consisting of 332 sets of target
structures, we observed better \Obj{} for Boltzmann sampling than for
uniform sampling (see Supp. Mat.~\ref{sec:validity}).  Notably,
\ourprog{} performs equally well in the presence of pseudoknots;
the difficulty rather lies in the computation of the \Obj{} function, since
free-energy minimization is costly in the presence of
pseudoknots~\citep{Sheikh2012} and good implementations are scarce.

%
Moreover, for all instances as well, the superiority of Boltzmann designs
remains unchallenged even after local optimizations, as shown by
Fig.~\ref{fig:benchmark-results} and Supp. Mat.~\ref{sec:improvements}
and \ref{sec:validity}. This observation is consistent with a
superior quality of the starting point for the greedy walk, probably leading
to better local minima of the \Obj{} function. However, it should be noted
that the greedy walk is based on the uniform sampling of
\RNAblueprint{}, and thus can be expected to partially level the
advantages of Boltzmann sampling. In future work, we hope to improve
this aspect by exploiting Boltzmann sampling during the
optimization run.

%\begin{table}[t]
%\centering
%\medskip
%\begin{tabular}{@{}>{\bf}l@{\quad}>{\tt}l@{\quad}@{\quad}c@{\quad}c@{\quad}c@{}}
%  &   \textbf{\textrm{Dataset}}   & {\bfseries\ourprog{}} & \textbf{\RNAblueprint} & \textbf{$\Delta$-Improvement}\\ \toprule
%  Seeds      & 2str & 21.69 ($\pm$4.37) & 37.74 ($\pm$6.45) & 16.05\\
%             & 3str & 18.10 ($\pm$3.98) & 30.49 ($\pm$5.41) & 12.39\\
%             & 4str & 20.15 ($\pm$3.83) & 32.32 ($\pm$5.23) & 12.17\\
%             \cline{2-5}
%             & PK60 & 12.12 ($\pm$3.04) & 19.81 ($\pm$4.24) & 7.69\\
%             & PK80 & 13.81 ($\pm$3.63) & 24.05 ($\pm$4.93) & 10.24\\
%             & LE80 & 10.54 ($\pm$2.77) & 17.80 ($\pm$4.17) & 7.26\\
%             \midrule
%
%  Optimized  & 2str & 5.42 ($\pm$1.27) & 7.95 ($\pm$1.75) & 2.52\\
%             & 3str & 5.10 ($\pm$1.10) & 7.04 ($\pm$1.52) & 1.94\\
%             & 4str & 9.17($\pm$1.51) & 13.25 ($\pm$2.13) & 4.17\\
%             \cline{2-5}
%             & PK60  & 3.84 ($\pm$1.22) & 6.45 ($\pm$1.84) & 2.60\\
%             & PK80* & 3.41 ($\pm$1.59) & 7.99 ($\pm$2.35) & 4.42\\
%             & LE80  & 3.59 ($\pm$1.11) & 5.88 ($\pm$1.59) & 2.28\\
%             \bottomrule
%\end{tabular}\\[1em]
%
%\caption{Comparison of the multi-stable design objective function
%  (Eq.~(\ref{eq:blueprintobjective}) of sequences sampled by
%  \ourprog{} vs.\ uniform samples for three benchmark sets; before
%  (seeds) and after optimization (optimized). We report the average
%  scores with standard deviations; smaller scores are better. The
%  final column reports our improvements due to Boltzmann sampling over
%  uniform sampling (as difference of mean scores).}
%\label{tab:benchmark-results}
%\end{table}

\begin{figure}
  {\includegraphics[width=.9\linewidth,trim={.2cm .3cm .2cm 5cm},clip]{Figs/statistics-overall}}
  \caption{Comparison of the \Obj{} (see
    Eq.~\eqref{eq:blueprintobjective}; smaller values $\to$ better
    designs) for sequences sampled by \ourprog{} and uniform sampling
    (\RNAblueprint) for six benchmark sets. For both sampling schemes,
    we show the MultiDefects of the sampled sequences, as well as
    results after further optimization by local search (``+ Opt.'').}
\label{fig:benchmark-results}
\end{figure}

\section*{Conclusion}
%%!!!!!!!!!!!!!!! consider to merge into discussion
%Going beyond our encoding of the base pair model energy features and the \GCb\% feature, the contributions provide more generality and allow expressing complex features of the
%sequences alone, e.g.\ rewarding or penalizing specific sequence
%motifs, as well as features depending on the target structures.
%Furthermore, constraints, which enforce or forbid features, are
%naturally expressed by assigning infinite penalties to invalid
%(partial) sequences (even if in implementations, explicit handling of constraints is advantageous). In particular, the %framework
%captures various RNA energy models, they can be handled efficiently (in the sense of FPT), as long as the dependencies of their additive energy contributions are bounded.

We introduced a novel approach, based on a general framework and efficient algorithms, to design RNA
sequences, while targeting very specific complex properties; in particular, we describe the targeting of the free energies of multiple target strucutures and the \GCb-content.
Our method combines an FPT stochastic sampling algorithm
with multi-dimensional Boltzmann sampling over distributions
controlled by expressive RNA energy models. Compared to the
previously best available sampling method (uniform sampling), this
approach generated significantly better seed sequences for instances
of an extensive multi-target design benchmark, including pseudoknots.
%
To substantiate our work additionally, we establish the $\#${\sf P}-hardness of uniform sampling;
from a complexity-theoretic point of view, this motivates the FPT, tree decomposition-based nature of our method.

The presented method enables new possibilities for the sequence generation
in the field of RNA sequence design, allowing to enforce additional
constraints, like \GCb\%, while controlling the energy of
multiple target structures. Thus it presents a major advance over
previously applied ad-hoc sampling and even efficient uniform sampling
strategies. We demonstrated the practical feasibility of such controlled sequence
generation, including the case of pseudoknotted structures, and studied its use for multi-target RNA design. Moreover,
our framework is extensible to include various more complex sequence
constraints, including mandatory/forbidden motifs at specific
positions, or anywhere in the designed sequences, by adapting formal language constructs
introduced by \citet{Zhou2013}. In future work, negative design principles could 
be explicitly supported at the generation stage, for instance by
penalizing a set of alternative helices/structures.
{\em Add RNASKetch to conclusion for more complex design scenario + need for generic constratins-based framework for flexibility}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
  YP and SW initiated the project and worked out the algorithms.
  % SW---I guess we don't need to be super-detailed, otherwise something like...
  %YP contributed the multi-dimensional Boltzmann sampling background and the \#P-hardness proof.
  %SW contributed the integration with the constraint network setting and its specialization to RNA design.
  YP and WW wrote the core software, while SW and SH contributed additional scripts. SH devised and conducted the computational experiments. All authors wrote, read, and approved the manuscript.
    % Text for this section \ldots
    % https://www.biomedcentral.com/getpublished/editorial-policies#authorship

\section*{Acknowledgements}
YP is supported
%jointly supported
by the
% French
{\em Agence
  Nationale de la Recherche} and the
%Austrian
FWF (ANR-14-CE34-0011; project RNALands).  SH is supported by the
German Federal Ministry of Education and Research (BMBF support code
031A538B; de.NBI: German Network for Bioinformatics Infrastructure)
and the
% acknowledged by
%financial support of the
Future and Emerging Technologies programme
% within the Seventh Framework Programme for Research of the
%European Commission, under
(FET-Open grant 323987; project RiboNets).
%
We thank Leonid Chindelevitch for suggesting a drastic optimization of our FPT algorithm based on stable sets combinatorics, and Arie Koster for practical recommendations on tree decompositions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{biblio}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

 \section*{Figures}
%   \begin{figure}[h!]
%   \caption{\csentence{Sample figure title.}
%       A short description of the figure content
%       should go here.}
%       \end{figure}
% 
% \begin{figure}[h!]
%   \caption{\csentence{Sample figure title.}
%       Figure legend text.}
%       \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
% \section*{Tables}
% \begin{table}[h!]
% \caption{Sample table title. This is where the description of the table should go.}
%       \begin{tabular}{cccc}
%         \hline
%            & B1  &B2   & B3\\ \hline
%         A1 & 0.1 & 0.2 & 0.3\\
%         A2 & ... & ..  & .\\
%         A3 & ..  & .   & .\\ \hline
%       \end{tabular}
% \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \section*{Additional Files}
%   \subsection*{Additional file 1 --- Sample additional file title}
%     Additional file descriptions text (including details of how to
%     view the file, if it is in a non-standard format or the file extension).  This might
%     refer to a multi-page table or a figure.
%
%  \subsection*{Additional file 2 --- Sample additional file title}
%    Additional file descriptions text.

\end{backmatter}
\end{document}